---
title: "Handling missing values!"
author: "jmpark"
output: 
  html_document :
    df_print : paged
---

### LIBRARY
```{r, message=FALSE, warning=FALSE}
library(knitr)
library(kableExtra)
library(tidyverse)
library(cluster)

```


### LOAD DATA
```{r, message=FALSE, warning=FALSE}
algae <- read.table('C:/Users/jeong/Desktop/데이터마이닝/Analysis.txt',
                    header=F,
                    dec='.',
                    col.names=c('season','size','speed','mxPH','mnO2','Cl','NO3',
                                'NH4','oPO4','PO4','Chla','a1','a2','a3','a4','a5','a6','a7'),
                    na.strings=c('XXXXXXX'))

algae<-algae[-c(62,199),]   # too many NA's
dim(algae)

```

### Missing values
  * mean (`mean`)
  * median (`median`)
  * simple linear regression using highly correlated variables (`cor`,`lm`)
  * closest distance (`cluster::daisy`)
```{r, message=FALSE, warning=FALSE}
nrow(algae[!complete.cases(algae),])
algae[!complete.cases(algae),]  

```  
</br>

#### 1. mean
```{r, message=FALSE, warning=FALSE}
summary(algae$mxPH)
algae[is.na(algae$mxPH),1:11] %>% 
  mutate(mxPH = cell_spec(mxPH, "html", color = ifelse(is.na(mxPH), "red", "black"))) %>%
  kable(format = "html", escape = F) %>%
  kable_styling("striped", full_width = F)
which(is.na(algae$mxPH))
mean(algae$mxPH,na.rm=T)
algae[48,'mxPH']<-mean(algae$mxPH,na.rm=T)
algae[48,1:11] %>% kable %>%  kable_styling(full_width = F)
summary(algae$mxPH)  

```  
</br>

#### 2. median
```{r, message=FALSE, warning=FALSE}
summary(algae$Chla)
algae[is.na(algae$Chla),1:11] %>% 
  mutate(Chla = cell_spec(Chla, "html", color = ifelse(is.na(Chla), "red", "black"))) %>%
  kable(format = "html", escape = F) %>%
  kable_styling("striped", full_width = F)
missing <- which(is.na(algae$Chla))
missing
algae[is.na(algae$Chla),'Chla']<-median(algae$Chla,na.rm=T)
algae[missing,1:11] %>% kable %>%  kable_styling(full_width = F)
summary(algae$Chla)  

```
</br>

#### 3. simple linear regression with a highly correlated variable
  * Find a variable that has a high correlation with the missing variable
  * Fit a simple linear regression
```{r, message=FALSE, warning=FALSE, fig.align='center'}
summary(algae$PO4)
algae[is.na(algae$PO4),1:11] %>% 
  mutate(PO4 = cell_spec(PO4, "html", color = ifelse(is.na(PO4), "red", "black"))) %>%
  kable(format = "html", escape = F) %>%
  kable_styling("striped", full_width = F)
missing <- which(is.na(algae$PO4))
missing
colnames(algae)
sub_algae <- algae[,c(4:11)]
names(sub_algae)
round(cor(sub_algae, use = 'complete.obs'),2)
symnum(cor(sub_algae, use="complete.obs"))
# cor(PO4, oPO4) = 0.91 => highly correlated!!
plot(sub_algae)

# lm : used to fit linear models
lm1<-lm(PO4~oPO4,data=algae)
summary(lm1)
algae[is.na(algae$PO4),'PO4']<-lm1$coef[1] + lm1$coef[2] * algae$oPO4[is.na(algae$PO4)]

# make it to a function
# use `oPO4` to fill NA's of `PO4`
fillPO4<-function(oP){
  if(is.na(oP)) return(NA)
  else return (oP * lm1$coef[2] + lm1$coef[1])
}
algae[is.na(algae$PO4),'PO4'] <- sapply(algae[is.na(algae$PO4),'oPO4'],fillPO4)
algae[missing,1:11] %>% kable %>%  kable_styling(full_width = F)
summary(algae$PO4)

```
</br>

#### 4. closest distance
  * use observations that are similar to the observations that include the missing value
```{r, message=FALSE, warning=FALSE}
summary(algae$mnO2)
algae[is.na(algae$mnO2),1:11] %>% 
  mutate(mnO2 = cell_spec(mnO2, "html", color = ifelse(is.na(mnO2), "red", "black"))) %>%
  kable(format = "html", escape = F) %>%
  kable_styling("striped", full_width = F)
missing <- which(is.na(algae$mnO2))
missing
algae1<-algae[,1:11]
# computes all the pairwise dissimilarities(distances) between observations in the data set
dist.mtx<-as.matrix(daisy(algae1,stand=T))
dim(dist.mtx)
# 10 closest observations
sort(dist.mtx[is.na(algae$mnO2),])[2:11]
aa<-as.integer(names(sort(dist.mtx[is.na(algae$mnO2),])[2:11]))
aa
algae[aa,'mnO2']
mean(algae[aa,'mnO2'])
algae[is.na(algae$mnO2),'mnO2'] <- mean(algae[aa,'mnO2'])
algae[missing,1:11] %>% kable %>%  kable_styling(full_width = F)
summary(algae$mnO2)

```


  * if `NA` is a numeric variable -> median
  * if `NA` is a factor -> voting(majority)
  * else -> change the variable to a factor
```{r, message=FALSE, warning=FALSE}
central.value <- function(x) {
  if (is.numeric(x)) median(x,na.rm=T)
  else if (is.factor(x)) levels(x)[which.max(table(x))]
  else {
    f <- as.factor(x)
    levels(f)[which.max(table(f))]
  }
}

```


```{r, message=FALSE, warning=FALSE}
# find rows that have NA values
# as.integer(names(sort(dist.mtx[r,])[2:11]))
# find 10 observations that are closest

for(r in which(!complete.cases(algae))) 
  algae[r,which(is.na(algae[r,]))] <- apply(
    data.frame(algae[c(as.integer(names(sort(dist.mtx[r,])[2:11]))), 
                     which(is.na(algae[r,]))]), 
    2, central.value)
summary(algae)
# NO missing values!!

```

































